Conflict-Free Vertex Coloring Of Planar Graphs

### 2017-04-03 paper/paper.tex
Making progress towards six pages

@@

A dominating set is a subset $D$ of $V$ such that every vertex not in $D$ is adjacent to at least one vertex in $D$. The domination number of $G$, denoted $\gamma(G)$, is the size of a minimum dominating set of G. The \emph{conflict-free domination number} for some $k$, denoted $\gamma_{CF}^k(G)$, is the minimum number of vertices that have to be colored in a conflict-free k-coloring of $G$. An example of a dominating set is $\{3, 5\}$ of graph $G_1$ where $\gamma(G_1) = 2$.

@@

@@

Abel et al. give a criterion that guarantees k-colorability which is stated in theorem \ref{thm:criterion}. In the proof of this theorem, found in \cite{abel2017three}, the authors prove that their heuristic will always generate a k-colorable graph when the specified criterion is met. To illustrate this heuristic, we will generate a graph that fits the specified criterion and apply the iterated elimination of distance-3-sets algorithm defined as algorithm \ref{alg:elimination}.

@@

\begin{theorem} \label{thm:four}
Every loopless planar graph admits a vertex coloring with at most four distinct colors
\end{theorem}

@@

\begin{theorem} \label{thm:four-cf}
Every loopless planar graph admits a conflict-free coloring with at most three distinct colors
\end{theorem}

@@

\subsection{Outerplanar Graphs}

\subsection{Planar Graphs}
As stated in theorem \ref{thm:four-cf} in section \ref{sec:bounds}, all planar graphs are 3-colorable.

@@

There are many applications of the CFCP that benefit from the recent research into the field. It is easier to minimize frequency assignment for wireless networks and model RFID networks. Current research focuses on finding bounds and properties on specific cases such as planar graphs, interval graphs, types of hypergraphs, and more. There is also recent interest in variations of the CFCP such as requiring \emph{another} vertex (not the selected vertex) to be colored within the neighborhood of a selected vertex. This allows for applications such as guiding a specific robot to other locations (i.e. the robot and its destination would both be a distinct unique color).

>>>>

\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\Input{\item[\algorithmicinput]}

@@

A dominating set is a subset $D$ of $V$ such that every vertex not in $D$ is adjacent to at least one vertex in $D$. The domination number of $G$, denoted $\gamma(G)$, is the size of a minimum dominating set of G. The \emph{conflict-free domination number} for some $k$, denoted $\gamma_{CF}^k(G)$, is the minimum number of vertices that have to be colored in a conflict-free k-coloring of $G$. An example of a dominating set is $\{3, 5\}$ of graph $G_1$ where $\gamma(G_1) = 2$. The \emph{k-conflict-free dominating set} problem asks if a given k-coloring of a graph colors the minimum amount of vertices needed to have a coloring that utilizes $k$ colors.

@@

It is often desirable and necessary to use approximation algorithms, known as heuristics, to solve complex problems like the VCP and the CFCP. In our case, heuristics are polynomial-time algorithms that give good, but not optimal, solutions to our optimization problems. A polynomial-time approximation scheme (PTAS) is any polynomial-time approximation algorithm that takes an instance of an optimization problem and a parameter $\epsilon > 0$ and produces a solution that is within a factor $(1 + \epsilon)$ (for minimization problems) of being optimal.

@@

Abel et al. give a criterion that guarantees k-colorability which is stated in theorem \ref{thm:criterion}. In the proof of this theorem, found in \cite{abel2017three}, the authors prove that their heuristic will always generate a k-colorable graph when the specified criterion is met. To illustrate this heuristic, we will generate a graph that fits the specified criterion and apply the iterated elimination of distance-3-sets algorithm defined in algorithm \ref{alg:elimination}.

@@

\begin{theorem} \label{thm:four}
Every loopless planar graph admits a vertex coloring with at most four distinct colors.
\end{theorem}

@@

\begin{theorem} \label{thm:four-cf}
Every loopless planar graph admits a conflict-free coloring with at most three distinct colors.
\end{theorem}

@@

\subsection{One or Two Colors}
\todo[inline]{To do still....}

\subsection{Three or More Colors}
Abel et. al \cite{abel2017three} present specific results for 3-colorable graphs. They examine minimizing the number of colored vertices of 3-colorable graphs. They prove theorems \ref{thm:approximate-cfds} and \ref{thm:outerplanar-cfdn} by using a polynomial-time algorithm that transforms a dominating set $D$ of a planar graph $G$ into a conflict-free k-coloring of $G$ where only the vertices of $D$ are colored. This algorithm is defined in algorithm \ref{alg:cf-dominating}.

\begin{algorithm}
\caption{Conflict-Free Coloring via Dominating Set} \label{alg:cf-dominating}
\begin{algorithmic}[1]
\Input A planar graph $G = (V, E)$
\State Find a dominating set of $G$
\While{$V \setminus D \neq \emptyset$}
	\ForAll{$v \in V \setminus D$}
		\State Pick a vertex $u \in D$
		\State Contract the edge $\{u,v\}$ towards $u$
	\EndFor
\EndWhile
\State Find a proper vertex coloring of $G$
\end{algorithmic}
\end{algorithm}

%To prove theorem \ref{thm:approximate-cfds}, we use a reduction from \emph{k-coloring} to \emph{conflict-free k-coloring}. This reduction is out of the scope of this paper and can be found in \cite{abel2017three}. The reduction preserves the planarity of its input graph, so the resulting graph is still planar. We also know that 3-coloring a graph is NP-complete \cite{sharma2012new}.

Because the planar minimum dominating set problem is NP-hard \cite{garey2002computers}, we know that the conflict-free dominating set problem is NP-hard because we are finding a minimum dominating set. It follows that the corresponding decision problem, the k-conflict-free dominating set problem, is NP-complete.

\begin{theorem} \label{thm:approximate-cfds}
For $k \geq 4$, the k-conflict-free dominating set problem is NP-complete and $\gamma_{CF}^k(G) = \gamma(G)$ for a planar graph $G$. Also, there exists a PTAS for estimating $\gamma_{CF}^k(G)$.
\end{theorem}

We get a PTAS for the conflict-free domination number by applying algorithm \ref{alg:cf-dominating} and finding the dominating set using the PTAS defined by Baker and Hill \cite{baker1994approximation}. As the input graph $G$ is planar, the resulting graph $G'$ is also planar because $G'$ is a minor of $G$. Thus, it can be colored with 4 colors by \ref{thm:four}. We then color $G$ with the colors of $G'$, leaving any vertex in $G$ but not in $G'$ uncolored. This leaves us with a conflict-free 4-coloring of $G$. Note that it has a minimal number of vertices colored because we used a PTAS to find the minimum dominating set.

\begin{theorem} \label{thm:outerplanar-cfdn}
If a graph $G$ is outerplanar, then $\gamma_{CF}^k(G) = \gamma(G)$. Also, there exists a linear-time algorithm to compute $\gamma_{CF}^k(G)$.
\end{theorem}

@@

There are many applications of the CFCP that benefit from the recent research into the field. It is easier to minimize frequency assignment for wireless networks and model RFID networks. Current research focuses on finding bounds and properties on specific cases such as planar graphs, interval graphs, types of hypergraphs, and more.

There is also recent interest in variations of the CFCP such as requiring \emph{another} vertex (not the selected vertex) to be colored within the neighborhood of a selected vertex. This allows for applications such as guiding a specific robot to other locations (i.e. the robot and its destination would both be a distinct unique color).
